---
title: "Final Assignment - Practical Machine Learning"
author: "Rotem Assouline"
date: "2023-12-30"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction 

In this project we analyse data from wearable devices in order to predict the manner in which participants preformed a dumbbell exercise. We load the dataset in perform a bit of preliminary processing:

```{r load}
library(caret)
library(dplyr)
set.seed(123)

# read the file
pml <- read.csv('pml-training.csv')

names(pml) <- gsub('picth','pitch',names(pml))

# change types of some columns
pml$classe <- as.factor(pml$classe) 
pml$user_name <- as.factor(pml$user_name)
pml$new_window <- pml$new_window == 'yes'
tonum <- (sapply(pml,class) %in% c('character','integer')) & (names(pml) != 'cvtd_timestamp')
pml[,tonum] <- lapply(pml[,tonum],as.numeric) 

# remove columns with missing values
pml <- pml[,colSums(is.na(pml)) == 0] 

str(pml)
```

We shall not use the first 7 columns (index, timestamp columns and window number).

```{r removecols}
pml <- pml[,8:dim(pml)[2]]
```

Next, we fit a random forest model to predict the outcome variable. We set the number of trees in the forest to 5, and perform 5-fold cross-validation.

```{r fit, cache = TRUE}
tC <- trainControl(method="cv", number=5, savePredictions = TRUE)
rfmod <- train(classe~.,
                data = pml,
                method = "rf",
                ntree = 5,
                trainControl = tC)
rfmod
```

Let us take a look at the confusion table and the in-sample error rate:

```{r check}
pred <- predict.train(rfmod,pml)
table(pred,pml$classe)
mean(pred == pml$classe)
```